A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-02-13 sections/Overview.tex
Declarative subtyping description

The syntax of this declarative system is shown in Figure~\ref{fig:decl:syntax}.

@@

In Figures~\ref{fig:decl:sub} and \ref{fig:decl:typing},
we give the subtyping and typing relations for the declarative system.

\paragraph{Overlapping rules}
\jimmy{Move to the Alg. Section}
Some of the declarative rules overlaps with each other.
Declarative subtyping rules $\mathtt{{\le}\forall L}$ and $\mathtt{{\le}\forall R}$
both match the conclusion $\Psi\vdash \forall a. A \le \forall a. B$.
In such case, choose $\mathtt{{\le}\forall R}$ first is always better,
since we introduce the type variable $a$ to the context earlier,
which gives more flexibility on the choice of $\tau$.

>>>>

The syntax of this declarative system is shown in Figure~\ref{fig:decl:syntax}.
A declarative type could be a unit type 1, a type variable $a$,
a universal quantification $\forall a. A$ or a function type $A \to B$.
Nested universal quantifiers are allowed for types,
but monotypes do not have any universal quantifier.
Context $\Psi$ is just a collection of type variable declarations.

@@

In Figures~\ref{fig:decl:sub} and \ref{fig:decl:typing},
we give the subtyping and typing relations for the declarative system,
which is exactly the same as DK's.
The subtyping relations originate from Odersky and La\"ufer~\cite{odersky1996putting},
and is adopted by DK as a part of their typing system.

\paragraph{Declarative Subtyping}
For implicit polymorphic type systems,
subtyping relation compares the degree of polymorphism of types.
In short, if a polymorphic type $A$ always succeed to instantiate to any instantiation of $B$,
then A is ``at least as polymorphic as'' $B$,
or we just say that $A$ is ``more polymorphic than'' $B$, written as $A \le B$.

Subtyping rules $\mathtt{{\le}Var}$, $\mathtt{{\le}Unit}$ and $\mathtt{{\le}{\to}}$
handle simple cases where no universal quantifiers get involved.
The subtyping rule for function types $\mathtt{{\le}{\to}}$ is standard,
being covariant on the return type and contravariant on the argument type.
Rule $\mathtt{{\le}\forall L}$ states that if $A$ is a subtype of $B$
under the context ($\Psi, a$), where $a$ is fresh in $A$, then $A\le\all B$.
Intuitively, if $A$ is more general than $\all B$ (where the universal quantifier
already indicates that $\all B$ is a general type),
then $A$ must instantiate to $[\tau/a]B$ for every $\tau$.

The most interesting rule is $\mathtt{{\le}\forall R}$.
If any instantiation of $\all A$, $[\tau/a]A$, is a subtype of $B$,
then $\all A \le B$.
The monotype $\tau$ we used to instantiation is \emph{guessed} for the declarative rule,
but the algorithmic system should be able to compute the guess properly.
The fact that $\tau$ is a monotype rules out the possibility of
polymorphic (or impredicative) instantiation.
Peyton Jones et al.~\cite{} also has such restriction of
predicative instantiation in their type system.
Both systems are adopt by several practical programming languages.

\paragraph{Declarative Typing}
\jimmy{TODO decl. typing description}

\paragraph{Overlapping Rules}
\jimmy{Move to the Alg. Section}
Some of the declarative rules overlaps with each other.
Declarative subtyping rules $\mathtt{{\le}\forall L}$ and $\mathtt{{\le}\forall R}$
both match the conclusion $\Psi\vdash \all A \le \all B$.
In such case, choose $\mathtt{{\le}\forall R}$ first is always better,
since we introduce the type variable $a$ to the context earlier,
which gives more flexibility on the choice of $\tau$.
